---
title: "Homework 8"
author: "Kimberly McNeil"
date: "3/27/2026"
output:
  html_document:
    code_folding: show
linkcolor: blue
urlcolor: blue
---

```{r setup, include=FALSE, echo=TRUE, warning=FALSE, message=FALSE}
# remember warning=FALSE and message=FALSE only when done
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(Lock5Data)
library(dplyr)
```


# Numeric Data Analysis

## BMI Calculation

The following table details the first six rows for Sex, Weight, Height, and the newly added BMI from the Student Survey dataset:^[Question 1 Complete]

```{r}
bmi_data = StudentSurvey %>%
  mutate(BMI = round(701 * (Weight / Height^2), 1)) %>%
  select(Sex, Weight, Height, BMI)

kable(head(bmi_data), caption = "First 6 Rows: Sex, Weight, Height, and BMI")
```

## Binning BMI into categories

The following table shows the same table as before, but with labels for BMIs in certain ranges (Underweight: Below 18.5, Healthy Weight: 18.5–24.9, Overweight: 25.0–29.9, and Obese: 30.0 or higher):^[Question 2 Complete]

```{r}
bmi_data_categories = bmi_data %>%
  mutate(
    BMI_Category = cut(
      BMI,
      breaks = c(-Inf, 18.5, 25.0, 30.0, Inf),
      labels = c("Underweight", "Healthy Weight", "Overweight", "Obese"),
      right  = FALSE # makes intervals closed on the left instead
    )
  )

kable(head(bmi_data_categories), caption = "First 6 Rows: BMI with Category Labels")
```

## Highest and Lowest BMIs by Sex

The following table summarizes the highest and lowest BMIs for male and female students:^[Question 3 Complete]

```{r bmi-extremes}
bmi_ends = bmi_data %>%
  group_by(Sex) %>%
  summarise(
    Min_BMI = min(BMI, na.rm = TRUE),
    Max_BMI = max(BMI, na.rm = TRUE),
    .groups = "drop"
  )

kable(bmi_ends, caption = "Minimum and Maximum BMI by Sex")
```

## Greater of each SAT scores

The following table shows the first six rows of the students' Verbal and Math SAT scores along with a new column displaying which of the two is greater:^[Question 4 Complete]

```{r}
sat_greatest = StudentSurvey %>%
  mutate(greater_SAT = pmax(VerbalSAT, MathSAT, na.rm = TRUE))  %>%
  select(VerbalSAT, MathSAT, greater_SAT)

kable(head(sat_greatest), caption = "First 6 Rows: Verbal SAT, Math SAT, and the Greater of the Two")
```

## Z-scores for Total SAT Scores

The following table shows the first six rows of the students' total SAT scores along with the new column calculating their z-scores:^[Question 5 Complete]

```{r}
sat_z_scores = StudentSurvey %>%
  mutate(
    z_score = (SAT - mean(SAT, na.rm = TRUE)) / sd(SAT, na.rm = TRUE),
    z_score = round(z_score, 3)
  ) %>%
  select(SAT, z_score)

kable(head(sat_z_scores), caption = "First 6 rows: Total SAT Score and its Z-score")
```
